---
title: "Statistical inference w4 report"
author: "Stephane Marceau"
subtitle: Part 1 Simulation exercice
output:
  pdf_document: default
job: Coursera Statistical inference course
---

## Summary
During this study, we will investigate the exponential distribution in R and compare it with the Central Limit Theorem.  
We will simulate its distribution by a getting 1000 means of 40 random exponential values. We will then, compare the sample mean and sample mean sd with the Normal distribution $\mathcal{N}(µ,\sigma/\sqrt(n)$ to verify the CLT. Finally, we will carry out a quantile to quantile check in order to visualize their matching.

## Simulation
We review an exponential mean random variable by simulating a thousand times the average of 40 exponentials distribution with lambda = 0.2.
```{r, fig.height=3.5}
library(ggplot2)
lambda<-0.2
mns <- sapply(1:1000,function(i) mean(rexp(40,lambda)))
sample_mean <- mean(mns)
sample_sd <- sd(mns)
mns_dat <- data.frame(x=mns, conf=mns>quantile(mns,probs = .025) & mns<quantile(mns,probs = .975))
g<-ggplot(data=mns_dat,aes(x=x))+geom_histogram(alpha = .2, binwidth=.5, color = "black",fill="blue", aes(y = ..density..)) 
g+ggtitle("sample mean of 40 exp")+theme(plot.title = element_text(hjust = 0.5))+ geom_vline(xintercept = mean(mns_dat$x), color="green")
```

From this simulation we get  
- a mean of `r mean(mns)` which is very close to the theoretical distribution mean of $µ=1/\lambda=5$  
- a standard deviation of `r sd(mns)` similar to $sd=\sigma/\sqrt(n)=`r 5/sqrt(40)`$

## plotting the Normal distribution  
As per the central limit theorem, $\bar x = \mathcal{N}(µ,\sigma/\sqrt(n)$.
Then, we plot the normal distribution $\mathcal{N}(5,5/\sqrt(40)$ on top of our simulation histogram. 
```{r, fig.height=3.5}
g<-g+stat_function(fun = dnorm,args=list(mean=5,sd=5/sqrt(40)))
g+ggtitle("sample mean of 40 exp vs normal distribution")+theme(plot.title = element_text(hjust = 0.5))
```

We see that both simulations and the normal distribution look very similar.

## plotting quantile to quantile chart
We use a qqplot chart to assess our simulation output with the normal distribution
```{r, fig.height=3.5}
pvals <- seq(.5, .99, by = .01)
d <- data.frame(n= qnorm(pvals,mean = 5,sd = 5/sqrt(40)),e=quantile(x=mns,probs = pvals), p = pvals)
g <- ggplot(d, aes(x= n, y = e))
g <- g + geom_abline(size = 2, col = "lightblue")
g <- g + geom_line(size = 2, col = "black")
g <- g + geom_vline(xintercept = qnorm(0.975,mean = 5,sd = 5/sqrt(40)))
g <- g + geom_hline(yintercept = quantile(x=mns,probs = 0.975))
g
```
Quantile to quantile plot shows that the sample means and the normal distribution $\mathcal{N}(µ,\sigma/\sqrt(n)$ are very close. We can conclude than the sample mean of 40 exp is approximately normal with $\mathcal{N}(5,5/\sqrt(40)$. 