---
title       : Statistical inference w4 report
subtitle    : Part 2 Basic Inferential Data Analysis
author      : Stephane Marceau
job         : Coursera Statistical inference course
output      : pdf_document
---

## Summary
In this project, we're going to analyze the ToothGrowth data in the R datasets package.
First we will look at the data table and print out a data frame summary.
Then we will visualize the growth of guinea pig tooth with regards to supp and vitamin C dose.
Finally, we will compare tooth growth due to supp effect and then tooth growth due to vitamin C dose effect using the hypothesis testing.

## Printing out data table head
```{r}
data(ToothGrowth)
head(ToothGrowth)
```
## Summarising data
```{r}
summary(ToothGrowth)
```

## Plotting the data
### Reviewing tooth growth by supp
```{r, fig.height=3.5}
library(ggplot2)
g <- ggplot(ToothGrowth, aes(x = supp, y = len, group = factor(dose)))
g <- g  + geom_point(size =10, pch = 21, aes(fill = dose), alpha = .5)+ggtitle("Guinea pig tooth growth by delivery method")
g+theme(plot.title = element_text(hjust = 0.5))
```

### Reviewing tooth growth by dose
```{r, fig.height=3.5}
g <- ggplot(ToothGrowth, aes(x = dose, y = len, group=factor(supp)))
g <- g  + geom_point(size =10, pch = 21, aes(fill = supp), alpha = .5)+ggtitle("Guinea pig tooth growth by vitamin C dose")
g+theme(plot.title = element_text(hjust = 0.5))
```

## Spliting data into groups
```{r}
# grouping by supp
vc_group<-ToothGrowth[ToothGrowth$supp=="VC",]
oj_group<-ToothGrowth[ToothGrowth$supp=="OJ",]

# grouping by dose
dose_group1<-ToothGrowth[ToothGrowth$dose==.5,]
dose_group2<-ToothGrowth[ToothGrowth$dose==1,]
dose_group3<-ToothGrowth[ToothGrowth$dose==2,]
```

## Comparing tooth growth by supp
**Testing $H_o: µ_{vc} = µ_{oj}$ vs $H_a: µ_{vc}< µ_{oj}$**
Using t.test function with two variables not paired
```{r}
t.test(x = vc_group$len,y = oj_group$len,mu = 0,paired = FALSE, var.equal = TRUE, alternative = "less")
```

p-value is below 5%. We reject the null hypothesis with a 95% confidence interval.
Using orange juice improves tooth growth compared to using Vitamin C.

## Comparing tooth growth by dose
**Testing $H_0: µ_1 = µ_2$ vs $H_a: µ_1 < µ_2$**
Using t.test function with two variables not paired
```{r}
t.test(x = dose_group1$len,y = dose_group2$len,mu = 0,paired = FALSE, var.equal = FALSE,alternative = "less")
```

p-value is significantly lower than 5%, thus we reject the null hypothesis.
Group 2 tooth growth is greater than group 1. A dose of 1mg/day improves tooth growth.

**Testing $H_0: µ_2 = µ_3$ vs $H_a: µ_2 < µ_3$**
Using t.test function with two variables not paired
```{r}
t.test(x = dose_group2$len,y = dose_group3$len,mu = 0,paired = FALSE, var.equal = FALSE,alternative = "less")
```

p-value is significantly lower than 5%, thus we reject the null hypothesis.
Group 3 tooth growth is greater than group 2. A dose of 2mg/day improves tooth growth.

## Conclusion
Based on the sample data ToothGrowth, we  conclude that tooth growth if not affected by the supp factor, but rather by the dose of Vitamin C received by the Guinea Pigs.